---
title: "US Naming Trends"
output: html_notebook
---

Full baby name data provided by the SSA by state. For the source data see [here](https://www.ssa.gov/oact/babynames/limits.html).

## Setup

Choose a method for reading file data. The names by state file contain over 6 million records.

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(dplyr)
library(ggplot2)
library(readr)

babynames <- read_csv("app/namesbystate.csv", col_types = cols(sex = col_character()))
#babynames <- vroom::vroom("app/namesbystate.csv")
#babynames <- data.table::fread("app/namesbystate.csv")
```

## Aggregate data by name

Identify the top 1000 names.

```{r}
topNames <- babynames %>%
  filter(year >= 1996) %>%  
  group_by(name, sex) %>%
  summarize(n = as.numeric(sum(n))) %>%
  filter(n > 1000) %>%
  select(name, sex)

filteredNames <- babynames %>%
  filter(year >= 1996) %>%
  inner_join(topNames)

yearlyNames <- filteredNames %>%
  group_by(year, name, sex) %>%
  summarize(n = as.numeric(sum(n)))
```

## Most popular names (1996)

Identify the top 3 male and female names from 1986. Visualize the popularity trend over time.

```{r}
topNames1996 <- yearlyNames %>%
  filter(year == 1996) %>%
  group_by(name, sex) %>%
  summarize(n = sum(n)) %>%
  group_by(sex) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 3) %>%
  arrange(sex, rank) %>%
  select(name, sex, rank)

topNames1996Yearly <- yearlyNames %>%
  inner_join(topNames1996)

ggplot(topNames1996Yearly, aes(year, n, color=name)) +
  facet_grid(~sex) +
  geom_line() +
  ggtitle("Most Popular Names of 1996")
```

## Most popular names (2020)

Identify the top 3 male and female names from 2014. Visualize the popularity trend over time.

```{r}
topNames2020 <- yearlyNames %>%
  filter(year == 2020) %>%
  group_by(name, sex) %>%
  summarize(n = sum(n)) %>%
  group_by(sex) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank <= 3) %>%
  arrange(sex, rank) %>%
  select(name, sex, rank)

topNames2020Yearly <- yearlyNames %>%
  inner_join(topNames2020)

ggplot(topNames2020Yearly, aes(year, n, color=name)) +
  facet_grid(~sex) +
  geom_line() +
  ggtitle("Most Popular Names of 2020")
```

## Shared names

Visualize the most popular names that are shared by both males and females.

```{r}
sharedName <- babynames %>%
  mutate(male = ifelse(sex == "M", n, 0), female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
  summarize(Male = as.numeric(sum(male)), 
            Female = as.numeric(sum(female)),
            n = as.numeric(sum(n)),
            AvgYear = round(as.numeric(sum(year * n) / sum(n)),0)) %>%
  filter(Male > 50000 & Female > 50000)

library(rbokeh)
figure(width = NULL, height = NULL, 
       xlab = "Log10 Number of Males", 
       ylab = "Log10 Number of Females",
       title = "Top shared names (1880 - 2020)") %>%
  ly_points(log10(Male), log10(Female), data = sharedName,
            color = AvgYear, size = scale(sqrt(n)),
            hover = list(name, Male, Female, AvgYear), legend = FALSE) %>%
  ly_abline(0, 1, color = "lightgrey")
```
